From Reversible Programs to \\ Univalent Universes and Back

### 2017-07-03 upi.tex
state theorems

\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}
\usepackage[inline]{enumitem}
\usepackage{agda}

\newcommand{\byiso}[1]{{\leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\byisotwo}[1]{{\Leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\unitepl}{\texttt{unitepl}}
\newcommand{\unitipl}{\texttt{unitipl}}
\newcommand{\unitepr}{\texttt{unitepr}}
\newcommand{\unitipr}{\texttt{unitipr}}
\newcommand{\swap}{\textit{swap}}
\newcommand{\swapp}{\texttt{swapp}}
\newcommand{\assoclp}{\texttt{assoclp}}
\newcommand{\assocrp}{\texttt{assocrp}}
\newcommand{\unitetl}{\texttt{unitetl}}
\newcommand{\unititl}{\texttt{unititl}}
\newcommand{\unitetr}{\texttt{unitetr}}
\newcommand{\unititr}{\texttt{unititr}}
\newcommand{\swapt}{\texttt{swapt}}
\newcommand{\assoclt}{\texttt{assoclt}}
\newcommand{\assocrt}{\texttt{assocrt}}
\newcommand{\absorbr}{\texttt{absorbr}}
\newcommand{\absorbl}{\texttt{absorbl}}
\newcommand{\factorzr}{\texttt{factorzr}}
\newcommand{\factorzl}{\texttt{factorzl}}
\newcommand{\factor}{\texttt{factor}}
\newcommand{\distl}{\texttt{distl}}
\newcommand{\dist}{\texttt{dist}}
\newcommand{\factorl}{\texttt{factorl}}
\newcommand{\id}{\textit{id}}
\newcommand{\compc}[2]{#1 \circ #2}
\newcommand{\compcc}[2]{#1 \bullet #2}
\newcommand{\respcomp}[2]{#1 \odot #2}

\newcommand{\Typ}{\mathbf{Type}}
\newcommand{\alt}{~\mid~}
\newcommand{\patht}[1]{\textsc{PATHS}(#1,#1)}
\newcommand{\fpatht}[1]{\textsc{FREEPATHS}(#1,\Box)}
\newcommand{\fpathp}[2]{\textsc{freepath}~#1~#2}
\newcommand{\pathind}[2]{\textsc{pathind}~#1~#2}
\newcommand{\invc}[1]{!\;#1}
\newcommand{\evalone}[2]{eval(#1,#2)}
\newcommand{\evalbone}[2]{evalB(#1,#2)}
\newcommand{\reflp}{\textsc{refl}}
\newcommand{\notp}{\textsc{not}}
\newcommand{\gluep}{\textsc{glue}}
\newcommand{\reflh}{\mathit{refl}_{\sim}}
\newcommand{\symh}[1]{\mathit{sym}_{\sim}~#1}
\newcommand{\transh}[2]{\mathit{trans}_{\sim}~#1~#2}
\newcommand{\reflq}{\mathit{refl}_{\simeq}}
\newcommand{\symq}[1]{\mathit{sym}_{\simeq}~#1}
\newcommand{\transq}[2]{\mathit{trans}_{\simeq}~#1~#2}
\newcommand{\isequiv}[1]{\mathit{isequiv}(#1)}
\newcommand{\idc}{\mathit{id}_{\boolt}}
\newcommand{\swapc}{\mathit{swap}_{\boolt}}
\newcommand{\assocc}{\mathit{assoc}}
\newcommand{\invl}{\mathit{invl}}
\newcommand{\invr}{\mathit{invr}}
\newcommand{\invinv}{\mathit{inv}^2}
\newcommand{\idlc}{\mathit{idl}}
\newcommand{\idrc}{\mathit{idr}}
\newcommand{\swapswap}{\swapc^2}
\newcommand{\compsim}{\compc_{\isotwo}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\isotwo}{\Leftrightarrow}
\newcommand{\piso}{\multimapdotbothB~~}
\newcommand{\zt}{\mathbb{0}}
\newcommand{\ot}{\mathbb{1}}
\newcommand{\bt}{\mathbb{2}}
\newcommand{\fc}{\mathit{false}}
\newcommand{\tc}{\mathit{true}}
\newcommand{\boolt}{\mathbb{B}}
\newcommand{\univ}{\mathcal{U}}
\newcommand{\uzero}{\mathcal{U}_0}
\newcommand{\uone}{\mathcal{U}_1}
\newcommand{\Rule}[2]{
\makebox{
$\displaystyle
\frac{\begin{array}{l}#1\\\end{array}}
{\begin{array}{l}#2\\\end{array}}$}}
\newcommand{\proves}{\vdash}
\newcommand{\jdgg}[3]{#1 \proves #2 : #3}
\newcommand{\jdg}[2]{\proves #1 : #2}
\newcommand{\jdge}[3]{\proves #1 = #2 : #3}
%% codes
%% denotations

% TODO: give this a better name!
\newcommand{\bracket}[1]{\ensuremath{\{#1\}}}
\newcommand{\ptrunc}[1]{\ensuremath{\| #1 \|}}

\newcommand{\PiTwo}{\ensuremath{\Pi_{\mathbb{2}}}}

@@

>>>>

\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}
\usepackage[inline]{enumitem}
\usepackage{agda}
\usepackage{stmaryrd}

\newcommand{\byiso}[1]{{\leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\byisotwo}[1]{{\Leftrightarrow}{\langle} ~#1~ \rangle}
\newcommand{\unitepl}{\texttt{unitepl}}
\newcommand{\unitipl}{\texttt{unitipl}}
\newcommand{\unitepr}{\texttt{unitepr}}
\newcommand{\unitipr}{\texttt{unitipr}}
\newcommand{\swap}{\textit{swap}}
\newcommand{\swapp}{\texttt{swapp}}
\newcommand{\assoclp}{\texttt{assoclp}}
\newcommand{\assocrp}{\texttt{assocrp}}
\newcommand{\unitetl}{\texttt{unitetl}}
\newcommand{\unititl}{\texttt{unititl}}
\newcommand{\unitetr}{\texttt{unitetr}}
\newcommand{\unititr}{\texttt{unititr}}
\newcommand{\swapt}{\texttt{swapt}}
\newcommand{\assoclt}{\texttt{assoclt}}
\newcommand{\assocrt}{\texttt{assocrt}}
\newcommand{\absorbr}{\texttt{absorbr}}
\newcommand{\absorbl}{\texttt{absorbl}}
\newcommand{\factorzr}{\texttt{factorzr}}
\newcommand{\factorzl}{\texttt{factorzl}}
\newcommand{\factor}{\texttt{factor}}
\newcommand{\distl}{\texttt{distl}}
\newcommand{\dist}{\texttt{dist}}
\newcommand{\factorl}{\texttt{factorl}}
\newcommand{\id}{\textit{id}}
\newcommand{\compc}[2]{#1 \circ #2}
\newcommand{\compcc}[2]{#1 \bullet #2}
\newcommand{\respcomp}[2]{#1 \odot #2}
\newcommand{\trunc}{\textit{trunc}}

\newcommand{\Typ}{\mathbf{Type}}
\newcommand{\alt}{~\mid~}
\newcommand{\patht}[1]{\textsc{PATHS}(#1,#1)}
\newcommand{\fpatht}[1]{\textsc{FREEPATHS}(#1,\Box)}
\newcommand{\fpathp}[2]{\textsc{freepath}~#1~#2}
\newcommand{\pathind}[2]{\textsc{pathind}~#1~#2}
\newcommand{\invc}[1]{!\;#1}
\newcommand{\evalone}[2]{eval(#1,#2)}
\newcommand{\evalbone}[2]{evalB(#1,#2)}
\newcommand{\reflp}{\textsc{refl}}
\newcommand{\notp}{\textsc{not}}
\newcommand{\gluep}{\textsc{glue}}
\newcommand{\reflh}{\mathit{refl}_{\sim}}
\newcommand{\symh}[1]{\mathit{sym}_{\sim}~#1}
\newcommand{\transh}[2]{\mathit{trans}_{\sim}~#1~#2}
\newcommand{\reflq}{\mathit{refl}_{\simeq}}
\newcommand{\symq}[1]{\mathit{sym}_{\simeq}~#1}
\newcommand{\transq}[2]{\mathit{trans}_{\simeq}~#1~#2}
\newcommand{\isequiv}[1]{\mathit{isequiv}(#1)}
\newcommand{\idc}{\mathit{id}_{\boolt}}
\newcommand{\swapc}{\mathit{swap}_{\boolt}}
\newcommand{\assocc}{\mathit{assoc}}
\newcommand{\invl}{\mathit{invl}}
\newcommand{\invr}{\mathit{invr}}
\newcommand{\invinv}{\mathit{inv}^2}
\newcommand{\idlc}{\mathit{idl}}
\newcommand{\idrc}{\mathit{idr}}
\newcommand{\swapswap}{\swapc^2}
\newcommand{\compsim}{\compc_{\isotwo}}
\newcommand{\iso}{\leftrightarrow}
\newcommand{\isotwo}{\Leftrightarrow}
\newcommand{\isothree}{\Lleftarrow \! \! \! \! \Rrightarrow}
\newcommand{\piso}{\multimapdotbothB~~}
\newcommand{\zt}{\mathbb{0}}
\newcommand{\ot}{\mathbb{1}}
\newcommand{\bt}{\mathbb{2}}
\newcommand{\fc}{\mathit{false}}
\newcommand{\tc}{\mathit{true}}
\newcommand{\boolt}{\mathbb{B}}
\newcommand{\univ}{\mathcal{U}}
\newcommand{\uzero}{\mathcal{U}_0}
\newcommand{\uone}{\mathcal{U}_1}
\newcommand{\Rule}[2]{
\makebox{
$\displaystyle
\frac{\begin{array}{l}#1\\\end{array}}
{\begin{array}{l}#2\\\end{array}}$}}
\newcommand{\proves}{\vdash}
\newcommand{\jdgg}[3]{#1 \proves #2 : #3}
\newcommand{\jdg}[2]{\proves #1 : #2}
\newcommand{\jdge}[3]{\proves #1 = #2 : #3}
%% codes
%% denotations

% TODO: give this a better name!
\newcommand{\bracket}[1]{\ensuremath{\{#1\}}}
\newcommand{\ptrunc}[1]{\ensuremath{\| #1 \|}}
\newcommand{\dbracket}[1]{\ensuremath{\llbracket \; #1 \; \rrbracket}}
\newcommand{\PiTwo}{\ensuremath{\Pi_{\mathbb{2}}}}

@@

We add a new level to $\{\PiTwo\}$ to prove the full correspondence of the first 2-levels of $\{\PiTwo\}$ and $\{\bt\}$:
\[\def\arraystretch{0.8}\begin{array}{l@{\quad}rclrl}
(\textit{3-combinators}) & \xi &::=& \trunc &:& \alpha \isothree \alpha
\end{array}\]

\begin{theorem}[Soundness and completeness for $\{\PiTwo\}$]
  There exist functions $\dbracket{ \_ }_n$ which map $n-$combinator in $\PiTwo$ to $n-$path in $\{\bt\}$,
  such that for all $n-$combinators $p, q$, $p \leftrightarrow_n q$ iff $\dbracket{ p }_n = \dbracket{ q }_n$.
\end{theorem}

\begin{theorem}[Soundness and completeness for $\{\bt\}$]
  There exist functions $\dbracket{ \_ }_n^{-1}$ which map $n-$path in $\{\bt\}$ to $n-$combinator in $\PiTwo$,
  such that for all $n-$path $p, q$, $p \equiv q$ iff $\dbracket{ p }_n^{-1} \isotwo \dbracket{ q }_n^{-1}$.
\end{theorem}
